---
title: "Untitled"
author: "Jason Torres"
date: "September 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 


```{r}
"%&%" <- function(a,b) paste0(a,b)

library("data.table")
library("dplyr")
library("Homo.sapiens")
library("ggplot2")
library(gridExtra)
library(grid)

server.dir <- "/Users/jtorres/FUSE/"

work.dir <- server.dir %&% "projects/mv-compare/"
input.dir <- work.dir %&% "input_files/"
output.dir <- work.dir %&% "output_files/"

known.genefile <- input.dir %&% "causal_genes3.txt"
metab.df <- fread(known.genefile)
plot.dir <- work.dir %&% "plots/"

mtxn.df <- fread(input.dir %&% "metaxcan_full_results_filtered_for_model_fit_and_pval_at_0.01.txt")

```


```{r}
# http://www.gettinggeneticsdone.com/2015/11/annotables-convert-gene-ids.html
#library("annotables")

get_ensid <- function(symb){
  #sub <- filter(grch37,symbol==symb,!(grepl("H",chr)))
  ensgene <- unique(filter(mtxn.df,gene_name==symb)$gene)
  if (length(ensgene)==0){
    ensgene <- unique(filter(grch37,symbol==symb,!(grepl("H",chr)))$ensgene)
  }
  return(ensgene)
}

get_symbol <- function(ensid){
  symb <- unique(filter(mtxn.df,gene==ensid)$gene_name)
  if (length(symb)==0){
    sub <- filter(grch37,ensgene==ensid)
    if (length(sub$symbol)==0){
      return(NA)
    } else{
      return(unique(sub$symbol)[1])
    }    
  } else{
    return(symb[1])
  }
}

get_metab_name <- function(metab){
  m <- filter(metab.df,MetaboID==metab)$metabolite
  if (length(m)==0){
    return(NA)
  } else{
    return(unique(m)[1])
  }
}

get_go_def <- function(goid){
  sub <- suppressMessages(select(Homo.sapiens,keys=goid,
                                 keytype="GO",columns="DEFINITION"))
  return(sub$DEFINITION[1])
}

ensid <- sapply(1:dim(metab.df)[1],function(i){
  symb <- metab.df$predicted_causal[i]
  unique(get_ensid(symb))
})

metab.df$ensid <- ensid


```


Build result data frame

```{r}


build_res_df <- function(){
  file.vec <- list.files(output.dir)
  out.df <- c()
  pb <- txtProgressBar(min=0,max=length(file.vec),style=3)
  for (i in 1:length(file.vec)){
    setTxtProgressBar(pb,i)
    f <- file.vec[i]
    build.df <- fread(output.dir%&%f)
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}


res.df <- build_res_df()

pb <- txtProgressBar(min=0,max=dim(res.df)[1],style=3)
res.df$symbol <- sapply(1:dim(res.df)[1],function(i){
  setTxtProgressBar(pb,i)
  ens <- res.df$dropped.gene[i]
  get_symbol(ens)
})

pb <- txtProgressBar(min=0,max=dim(res.df)[1],style=3)
res.df$metab <- sapply(1:dim(res.df)[1],function(i){
  setTxtProgressBar(pb,i)
  metab <- res.df$metabolite[i]
  get_metab_name(metab)
})

#pb <- txtProgressBar(min=0,max=dim(res.df)[1],style=3)
#res.df$metab <- sapply(1:dim(res.df)[1],function(i){
#  setTxtProgressBar(pb,i)
#  goid <- res.df$GO[i]
#  get_go_def(goid)
#})

```


```{r}

known <- c() 
pb <- txtProgressBar(min=0,max=dim(res.df)[1],style = 3)
for (i in 1:dim(res.df)[1]){
  setTxtProgressBar(pb,i)
  m <- res.df$metabolite[i]
  d <- res.df$dropped.gene[i]
  if (d =="full"){
    known <- append(known,NA)
  } else{
    sub <- filter(metab.df,MetaboID==m,ensid==d)
    if (dim(sub)[1]==0){
      known <- append(known,FALSE)
    } else{
      known <- append(known,TRUE)
    }
  }
}

res.df$known <- known
```

Add impact score 


```{r}


add_iscore <- function(){
  pairs <- unique(res.df$metabolite %&% "_" %&% res.df$GO)
  pb <- txtProgressBar(min=0,max=length(pairs),style=3)
  out.df <- c() 
  for (i in 1:length(pairs)){
    setTxtProgressBar(pb,i)
    vec <- strsplit(pairs[i],split="_")[[1]]
    mymetab <- vec[1]; mygo <- vec[2]
    sub.df <- filter(res.df,metabolite==mymetab,GO==mygo)
    b <- -log(filter(sub.df,dropped.gene=="full")$pval.adj,base=10)
    iscore <- b - (-log(sub.df$pval.adj,base=10))
    sub.df$iscore <- iscore  
    out.df <- rbind(out.df,sub.df)
  }
 return(out.df)
}

res.full.df <- add_iscore()

write.table(x=res.full.df,file=work.dir%&%"go-enrich-results_full.txt",
            sep="\t",quote=FALSE,row.names=FALSE)
```




# Profile GO results 


```{r}


#test.df <- filter(sig.df,metabolite=="M00606",GO=="GO:0008655")
#mymetab <- "M00606"; mygo <- "GO:0008655"

go_plot <- function(mymetab,mygo){
  plot.df <- filter(sig.df,metabolite==mymetab,GO==mygo)
  def <- get_go_def(mygo)
  #def <- substr(def,1,50) %&% "\n" %&% substr(def,51,100) %&% 
   # "\n" %&% substr(def,101,150) %&% "\n" %&% substr(def,151,200) %&% 
   # "..."
  def <- substr(def,1,100) %&% "\n" %&% substr(def,101,200) %&% 
    "\n" %&% substr(def,201,300) %&% "..."
  plt <- ggplot(data=plot.df,aes(x=dropped.gene,y=-log(pval.adj,base=10))) + 
    geom_hline(data=filter(plot.df,dropped.gene=="full"),
               aes(yintercept=-log(pval.adj,base=10)),
               color="firebrick4",linetype=3) + 
    geom_hline(yintercept=-log(0.05,base=10),
               color="firebrick4",linetype=2) +
    geom_point(data=filter(plot.df,dropped.gene!="full"),
               shape=23,size=4,aes(color=impact,fill=known)) +
    scale_y_continuous(limits=c(0,8),breaks=seq(0,8,1)) + 
    scale_color_manual(values=c("black","gray"),
                      name="Impact",labels=c("Increases \nGO enrichemnt",
                                             "Decreases \nGO enrichment")) + 
    scale_fill_brewer(type="qual",palette="Dark2", ######
                      name="Metabolite causal gene") + 
    theme_bw() + 
    theme(panel.grid.minor = element_blank(),
                       panel.grid.major = element_blank(),
                       axis.text.x = element_text(angle=90,size=6)) + 
    xlab("Dropped Gene") + 
    ylab(expression(paste(-log[10](`p-value`)))) + 
    geom_text(data=filter(plot.df,impact=="less.sig"),
              aes(x=dropped.gene,y=-log(pval.adj,base=10),
                  label=symbol),vjust = -2, nudge_x = 0.05,size=2) + 
    ggtitle("Metabolite: " %&% plot.df$metab[1] %&% " (" %&% 
              plot.df$metabolite[1] %&% ")" %&% 
              "\nPathway: " %&% plot.df$GO[1],
            subtitle = def)  
  plt
  ggsave(plot=plt,filename=plot.dir %&% mymetab %&% "-" %&% 
           gsub(":","_",mygo) %&% ".pdf",width=10,height=5)
}

go_plot(mymetab,mygo)

```

Accent, Dark2, Paired, Pastel1, Pastel2, Set1, Set2, Set3


Save plots 

```{r}

sig.df <- filter(res.full.df,pval.adj < 0.05)
write.table(x=sig.df,file=work.dir%&%"go-enrich-results_fdr05.txt",
            sep="\t",quote=FALSE,row.names=FALSE)

save_plots <- function(){
  pairs <- unique(sig.df$metabolite %&% "_" %&% sig.df$GO)
  pb <- txtProgressBar(min=0,max=length(pairs),style=3)
  for (i in 1:length(pairs)){
    setTxtProgressBar(pb,i)
    vec <- strsplit(pairs[i],split="_")[[1]]
    mymetab <- vec[1]; mygo <- vec[2]
    go_plot(mymetab,mygo)
  }
}

save_plots()

```


# Analysis significant results 

```{r}

res.full.df$iscore.scale <- scale(res.full.df$iscore)[1:dim(res.full.df)[1]]

sig.df <- filter(res.full.df,pval.adj < 0.05)

known.df <- filter(sig.df,known=="TRUE")
nov.df <- filter(sig.df,known=="FALSE")

qqnorm(res.full.df$iscore.scale)
qqline(res.full.df$iscore.scale,col="red")

qqnorm(nov.df$iscore.scale)
qqline(nov.df$iscore.scale,col="red")

qqnorm(known.df$iscore.scale)
qqline(known.df$iscore.scale,col="red")

shapiro.test(sample(res.full.df$iscore.scale,5000)) # Violates normality 
shapiro.test(sample(sig.df$iscore.scale)) # Violates normality 

ggplot(data=filter(sig.df,dropped.gene!="full"),aes(x=known,y=iscore.scale)) + 
         geom_jitter(position=position_jitter(0.2)) + 
  stat_summary(fun.data=mean_sdl, mult=1, 
               geom="pointrange", color="red")


tt <- t.test(nov.df$iscore.scale,known.df$iscore.scale)

wilcox.test(iscore.scale ~ known, data=sig.df) 

```

	Wilcoxon rank sum test with continuity correction

data:  iscore.scale by known
W = 9779.5, p-value = 0.0003349
alternative hypothesis: true location shift is not equal to 0



# Identity genes of interest 


```{r}

# Number of Metabolites with causal genes 
length(unique(metab.df$MetaboID)) # 74 
# Number of causal genes implicated with metabolites 
length(unique(metab.df$ensid)) # 93 

# Number of Metabolites with MetaXcan gene associations 
length(unique(mtxn.df$metabolite)) # 45 
# Number of genes associated with Metabolites 
length(unique(mtxn.df$gene)) # 736

# Proprotion of causal genes in MetaXcan gene set 
sum(unique(metab.df$ensid) %in% unique(mtxn.df$gene))/length(unique(metab.df$ensid)) # 0.44

# Proportion of metabolites with MetaXcan gene sets enriched for GO pathways 
sum(unique(mtxn.df$metabolite) %in% unique(sig.df$metabolite)) / length(unique(mtxn.df$metabolite)) # 0.31

# Number of enriched GO pathways 
length(unique(sig.df$GO)) # 72 
# Number of Known genes in enriched GO pathways 
length(unique(filter(sig.df,known==TRUE)$dropped.gene)) # 6 
unique(filter(sig.df,known==TRUE)$symbol)
# "TYMP"  "UGT1A1"  "SLC6A12" "SLC7A5"  "ACADS"   "SCAPER" 
# Number of enriched GO pathways corresponding to these 5 known genes 
length(unique(filter(sig.df,known==TRUE)$GO)) # 26
# Known genes that impact enrichments 
length(unique(filter(sig.df,known==TRUE,impact=="less.sig")$dropped.gene)) # TYMP
unique(filter(sig.df,known==TRUE,impact=="less.sig")$symbol)
# "TYMP" "UGT1A1"


# Number of 'unknown' genes in enriched GO pathways 
length(unique(filter(sig.df,known==FALSE)$dropped.gene)) # 188
# Number of 'unknown' genes in enriched GO pathways with impacts on enrichment
length(unique(filter(sig.df,known==FALSE,impact=="less.sig")$dropped.gene)) # 75



disc.df <- filter(sig.df,impact=="less.sig") %>% arrange(desc(iscore))
pairs <- unique(disc.df$metabolite %&% "-" %&% disc.df$GO)
#pairs <- sapply(1:length(pairs),function(i){
#  gsub(":","_",pairs[i])
#})

plot.dir <- "/Users/jtorres/Google Drive/Science/Projects/mv-compare/" %&% "plots/"

save_lesssig_plots <- function(){
  pairs <- unique(disc.df$metabolite %&% "_" %&% disc.df$GO)
  pb <- txtProgressBar(min=0,max=length(pairs),style=3)
  for (i in 1:length(pairs)){
    setTxtProgressBar(pb,i)
    vec <- strsplit(pairs[i],split="_")[[1]]
    mymetab <- vec[1]; mygo <- vec[2]
    go_plot(mymetab,mygo)
  }
}

save_lesssig_plots() 

```



```{r}

length(unique(filter(sig.df,impact=="less.sig",known==FALSE)$dropped.gene)) # 75 genes 
length(unique(filter(sig.df,known==FALSE)$dropped.gene)) # 188 genes 

#Potentially novel non-bystander genes 
length(unique(filter(sig.df,impact=="less.sig",known==FALSE)$dropped.gene))/
  length(unique(filter(sig.df,known==FALSE)$dropped.gene))
# 0.40 

#Potentially novel non-bystander genes (sd <= 1) 
length(unique(filter(sig.df,impact=="less.sig",known==FALSE,iscore.scale>=1)$dropped.gene)) / # 9 genes 
  length(unique(filter(sig.df,known==FALSE)$dropped.gene))
# 0.043 

length(unique(filter(sig.df,impact=="less.sig",known==FALSE,iscore.scale>=2)$dropped.gene))/ # 9 genes 
  length(unique(filter(sig.df,known==FALSE)$dropped.gene))
# 0.043, same as above 

cand.df <- filter(sig.df,impact=="less.sig",known==FALSE,iscore.scale>=2)
ggsave(plot=grid.table(cand.df),plot.dir %&% "cand-table.png",height=20,width=20)

```

Table plot of just enriched GO pathways 

```{r}

tab.df <- dplyr::select(sig.df,one_of("metabolite","metab","ontology","GO"))
tab.df <- tab.df[!duplicated(tab.df),]
tab.df$Def <- sapply(1:dim(tab.df)[1],function(i){
  goid <- tab.df$GO[i]
  def <- get_go_def(goid)
  def <- substr(def,1,100) %&% "\n" %&% substr(def,101,200) %&% "..."
  #"\n" %&% substr(def,101,150) %&% "..."
})

tab1.df <- tab.df[1:30,]
tab2.df <- tab.df[31:60,]
tab3.df <- tab.df[61:91,]

ggsave(plot=grid.table(tab1.df),plot.dir %&% "go-table1.png",height=20,width=20)
ggsave(plot=grid.table(tab2.df),plot.dir %&% "go-table2.png",height=20,width=20)
ggsave(plot=grid.table(tab3.df),plot.dir %&% "go-table3.png",height=20,width=20)


```

