---
title: "Untitled"
author: "Jason Torres"
date: "April 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup 


```{r}
"%&%" <- function(a,b) paste0(a,b)

library("data.table")
library("tidyverse")
library("gridExtra")
library("grid")

server.dir <- "/home/jason/science/servers/FUSE/"
server.dir2 <- "/home/jason/science/servers/FUSE5/"
work.dir <- server.dir %&% "projects/mv-compare/"
work.dir2 <- server.dir2 %&% "projects/mv-compare/"

input.dir <- work.dir %&% "input_files/"
output.dir <- work.dir %&% "output_files/"

#known.genefile <- input.dir %&% "causal_genes3.txt"
#known.df <- fread(known.genefile)
#metab.old.df <- fread(known.genefile)

plot.dir <- work.dir %&% "plots/"


mtxn.df <- fread(input.dir %&% "metaxcan_full_results_filtered_for_model_fit_and_pval_at_0.01.txt")
hmdb.df <- fread(work.dir %&% "hmdb/hmdb.txt")

```


Proximity annotation file 

```{r}

#oneMb.df <- fread(work.dir2 %&% "archive/all_sigGenes_apayne_291117_anno.txt")
#oneMb.genes <- filter(oneMb.df,pos_flag=="within-1MB-of-GWAS")$Gene %>% unique(.)

oneMb.df <- fread(work.dir2 %&% "all_sigGenes_apayne_171218_anno.txt")



```



```{r}

metab.df <- fread(input.dir %&% "metabolite_GWAS_genes_from-Anne.txt")
names(metab.df) <- c("gene_name","metabolite","MetaboID","predicted_causal")
# 58:   SLC22A2                   isobutyrylcarnitine   M33441          SLC22A1 # ask Anne for clarification


```



```{r}


mtxn.df$predicted_causal <- map(1:dim(mtxn.df)[1],function(i){
  m <- mtxn.df$metabolite[i]
  g <- mtxn.df$gene_name[i]
  val <- filter(metab.df,gene_name==g,predicted_causal!="unknown",MetaboID==m)$predicted_causal # g %in% 
  ifelse(length(val)>0,val,"-")
  }) %>% as.character(.)
  #as.logical(.)

mtxn.df$gwas.unknown <- map(1:dim(mtxn.df)[1],function(i){
  m <- mtxn.df$metabolite[i]
  g <- mtxn.df$gene_name[i]
  val <- filter(metab.df,gene_name==g,predicted_causal=="unknown",MetaboID==m)$gene_name # g %in% 
  ifelse(length(val)>0,val,"-")
  }) %>% as.character(.)
  #as.logical(.)

mtxn.df$hmdb <- map(mtxn.df$gene_name,function(g){g %in% hmdb.df$gene_name}) %>% 
  as.logical(.)

sub <- dplyr::select(mtxn.df,one_of("gene_name","metabolite","predicted_causal","gwas.unknown","hmdb"))
sub <- sub[!duplicated(sub),]
dim(sub) # 826 



dim(filter(sub,gene_name==predicted_causal,hmdb==FALSE))[1] # 1 predicted causal gene, SLC16A9 (for M15500 and M32412) not in HMDB
filter(sub,gene_name==predicted_causal,hmdb==TRUE)$gene_name %>% unique(.) %>% length(.) # 34 known genes in HMDB
dim(filter(sub,gene_name==gwas.unknown,hmdb==FALSE))[1] #  6 of the gwas unknown genes are NOT in HMDB
filter(sub,gene_name==gwas.unknown,hmdb==TRUE)$gene_name %>% unique(.) %>% length(.) # 4 of the gwas unknown genes are in HMDB


#sum(metab.df$predicted_causal %in% hmdb.df$gene_name)
#metab.df$predicted_causal[!(metab.df$predicted_causal%in% hmdb.df$gene_name)]


novel.df <- filter(sub,predicted_causal==FALSE,hmdb==TRUE)


filter(sub,gene_name!=predicted_causal) %>% dim(.) # 785 (704 genes) novel gene/metabolite pairs 
#filter(sub,gene_name!=predicted_causal)$gene_name %>% unique(.) %>% length(.)
filter(sub,gene_name!=predicted_causal,hmdb==TRUE) %>% dim(.) # 212 (189 genes) of the novel gene/metabolite pairs are in HMDB
# 212/785 = 27%, 189/704 = 0.268
#filter(sub,gene_name!=predicted_causal,hmdb==TRUE)$gene_name %>% unique(.) %>% length(.) # 
filter(sub,gene_name!=predicted_causal,gene_name==gwas.unknown,hmdb==TRUE) %>% dim(.) # 4 pairs of which (corresponding to 4 unique genes) are in HMDB and are gwas.unknown 
#      ANKH     M01564                -         ANKH TRUE
#     ACSM5     M15749                -        ACSM5 TRUE
#  RAB3GAP1     M20675                -     RAB3GAP1 TRUE
#     ABCC1     M37058                -        ABCC1 TRUE


# IGNORE FOR NOW 
## If I exclude genes within 1Mb of GWAS hits 
## ex <- filter(sub,gene_name!=predicted_causal,hmdb==TRUE,!(gene_name%in%oneMb.genes))$gene_name %>% unique(.) 
## length(ex)/704 # 0.1349 or 13.5% of novel genes are outside 1Mb of known met-QTL loci and are annotated to metabolic ## pathways


dim(filter(metab.df,predicted_causal!="unknown"))[1] # 61
dim(filter(metab.df,predicted_causal=="unknown"))[1] # 18


```


## Re-evaluation (first) 

826 Total significant gene-metabolite pairs 

41 gene-metabolite pairs involve genes that are predicted causal for the metabolite 
473 significant gene-metabolite pairs are within 1Mb of significant metQTLs, likely to be bystanders (from Anne's section of manuscript)
312 significant gene-metabolite pairs are not bystander associations and 
    do not involve predicted causal genes for the metabolite [could also be 312 = 826 - (41+473) ]

96 pairs are outside of 1Mb windows, not predicted causal for the metabolite, and are in hmdb; 
    corresponding to 83 unique genes 
Therefore, upwards of 30.8% (96/312) of novel genes-metabolite are biologically plausible 

```{r}

out.df <- c() 
for (i in 1:dim(sub)[1]){
  sub.df <- sub[i,]
  g <- sub.df$gene_name; m <- sub.df$metabolite; h <- sub.df$hmdb; c <- sub.df$predicted_causal
  check1 <- nchar(c) > 1 # is gene already a predicted causal gene for the metab? 
  check2 <- h == TRUE # is gene annotated to a metabolic pathway 
  # check 3 : is gene within 1Mb of GWAS sigal for the metabolite 
  check3 <- filter(oneMb.df,Metabo_id==m,Gene==g)$pos_flag == "within-1MB-of-GWAS" 
  if (check1==FALSE & check2==TRUE & check3==FALSE){
    out.df <- rbind(out.df,sub.df)
  }
}

filt.df <- out.df
filt.df %>% dim(.) # 96
96/312 # 0.3086817
filt.df$gene_name %>% unique(.) %>% length(.) # 83 genes 


```



```{r}


get_metab_annot <- function(query.gene){
  q.df <- filter(hmdb.df,gene_name==query.gene)
  if (dim(q.df)[1] > 0){
    go.vec <- strsplit(q.df$go_classifications,split="GO:\\d+",fixed=FALSE)[[1]]
    go.ids <- regmatches(q.df$go_classifications, gregexpr("GO:\\d+", q.df$go_classifications))[[1]]
    path.vec <- strsplit(q.df$pathways,split="map\\d{5}",fixed=FALSE)[[1]] %>% 
      map(.,function(s){strsplit(s,split="SMP\\d{5}",fixed=FALSE)[[1]]}) %>% unlist(.)
    return(list(go.ids,path.vec))    
  } 
  else{
    return(list(NA,NA))
  }
}

group_metab_assoc_genes <- function(metab){
  gene.vec <- filter(mtxn.df,metabolite==metab)$gene_name %>% unique(.)
  known.vec <- gene.vec[gene.vec %in% filter(metab.df,MetaboID==metab)$predicted_causal]
  novel.vec <- gene.vec[!(gene.vec %in% filter(metab.df,MetaboID==metab)$predicted_causal)]
  return(list(known.vec,novel.vec))
}

compare_gene_annotations <- function(gene1,gene2,annot.num){
  # annot.num = 1; use GO IDs 
  # annot.num = 2; use hmdb pathway names 
  g1 <- get_metab_annot(gene1)[[annot.num]]
  g2 <- get_metab_annot(gene2)[[annot.num]]
  overlap <- intersect(g1,g2)
  if (length(overlap)==0){
    return("None")
  } else{
    return(overlap)
  }
}

build_single_metab_df <- function(metab,alt=FALSE){
  l <- group_metab_assoc_genes(metab)  
  known.vec <- l[[1]]
  novel.vec <- l[[2]]

  out.df <- c()
  k.df <- data.frame(Gene=known.vec,
                     Group=rep("known",length(known.vec)),
                     go.shared.known=rep("-",length(known.vec)),
                     path.shared.known=rep("-",length(known.vec)),
                     stringsAsFactors = FALSE)
  out.df <- rbind(out.df,k.df)
  
  if (alt==TRUE){
    known.vec <- filter(metab.df,MetaboID==metab)$predicted_causal %>% unique(.)
  }
  known.goids <- map(known.vec,function(k){get_metab_annot(k)[[1]]}) %>% unlist(.) %>% unique(.) %>% na.omit(.)
  known.paths <- map(known.vec,function(k){get_metab_annot(k)[[2]]}) %>% unlist(.) %>% unique(.) %>% na.omit(.)  
  
  for (n in novel.vec){
    #print(n)
    go.ids <- get_metab_annot(n)[[1]]
    path.ids <- get_metab_annot(n)[[2]]
    
    path.eval <- any(path.ids %in% known.paths)==TRUE
    if (path.eval==FALSE & length(path.ids) > 0 & is.na(path.ids) ==FALSE){
      eval.vec <- c(unlist(map(path.ids,function(id){grepl(id,known.paths,fixed=TRUE)})),
                    unlist(map(known.paths,function(id){grepl(id,path.ids,fixed=TRUE)})))
      if (any(eval.vec)==TRUE){
        path.eval=TRUE
      }else{
        path.eval=FALSE
      }
    }
    
    build.df <- data.frame(Gene=n,
                     Group="novel",
                     go.shared.known=any(go.ids %in% known.goids)==TRUE,
                     path.shared.known=path.eval,
                     stringsAsFactors = FALSE)
    out.df <- rbind(out.df,build.df)
  }
  out.df <- data.frame(Metabolite=rep(metab,dim(out.df)[1]),out.df)
  return(out.df)
}

build_full_metab_df <- function(alt=FALSE){
  metab.vec <- unique(mtxn.df$metabolite)
  out.df <- c()
  pb <- txtProgressBar(min=0,max=length(metab.vec),style=3)
  for (i in 1:length(metab.vec)){
    setTxtProgressBar(pb,i)
    metab <- metab.vec[i]
    #print(metab)
    build.df <- build_single_metab_df(metab,alt)
    out.df <- rbind(out.df,build.df)
  }
  out.df$Metabolite <- as.character(out.df$Metabolite)
  return(out.df)
}

full.df <- build_full_metab_df(alt=TRUE)


```



## Table 2: Profile of novel Metaxcan genes, for those that are annotated to same metabolic pathway as known metabolite genes, to what extent are they significant?


```{r}


slice<-function(x,n) { # https://stackoverflow.com/questions/2434833/how-can-i-partition-a-vector
    N<-length(x);
    lapply(seq(1,N,n),function(i) x[i:min(i+n-1,N)])
}

make_table_2_df  <- function(){
  # Subset to novel genes annotated in hmdb pathways 
  sub.df <- filter(mtxn.df,predicted_causal=="-",hmdb==TRUE) %>% #,
                   #metabolite%in%filter(metab.df,predicted_causal!="unknown")$MetaboID) %>% 
    dplyr::select(one_of("gene","gene_name","metabolite","metabolite_name",
                     "Tissue","zscore","pvalue","FDR_studywide")) # 628 S-PrediXcan associations involve novel genes associated with metabolites with predicted causal genes 
 # t <- dplyr::select(sub.df,one_of("gene_name","metabolite")); t <- t[!duplicated(t),]
 # corresponding to 188 unique gene/metabolite pairs and 173 unique genes 
  pred.causal.gene <- c()
  shared.go.pathways <- c()
  shared.hmdb.pathways <- c() 
  pb <- txtProgressBar(min=0,max=dim(sub.df)[1],style=3)
  for (i in 1:dim(sub.df)[1]){
    #print(i)
    setTxtProgressBar(pb,i)
    g <- sub.df$gene_name[i]
    mid <- sub.df$metabolite[i]
    causal.genes.df <- filter(metab.df,MetaboID==mid)
    pcg <-  causal.genes.df$predicted_causal %>% paste(.,collapse=",")#map(causal.genes.df$predicted_causal,function(s){
      #vec <- strsplit(s,split=",")[[1]];sliced <- slice(vec,3)
      #map(sliced,function(vec){paste(vec,collapse=",")}) %>% unlist(.) %>% paste(.,collapse=",\n")
       #}) %>% as.character(.)
    query <- get_metab_annot(g)
    query.go <- query[[1]]
    query.hmdb <- query[[2]]
    reflists <-  ifelse(dim(causal.genes.df)[1]>0,map(1:length(causal.genes.df$predicted_causal), function(i){
      pc <- causal.genes.df$predicted_causal[i]
      metabs <- get_metab_annot(pc)
      return(metabs)
      }),NA)
    sgp <- ifelse(dim(causal.genes.df)[1]>0,map(1:length(reflists),function(i){ # 
      go.vec <- reflists[[i]][[1]]
      overlap <- go.vec[go.vec %in% query.go]
      if (length(overlap)==0){
        overlap <- NA
      }
      return(overlap)
    }) %>% unlist(.) %>% as.character(.)  %>% paste0(.,collapse=","),NA) 
    shp <- ifelse(dim(causal.genes.df)[1]>0,map(1:length(reflists),function(i){
      hmdb.vec <- reflists[[i]][[2]]
      overlap <- hmdb.vec[hmdb.vec %in% query.hmdb]
      if (length(overlap)==0){
        overlap <- NA
      }
      return(overlap)
    }) %>% unlist(.) %>% as.character(.) %>% paste0(.,collapse=","),NA)
    pred.causal.gene <- append(pred.causal.gene,pcg)
    shared.go.pathways <- append(shared.go.pathways,sgp)
    shared.hmdb.pathways <- append(shared.hmdb.pathways,shp)  
  }
  out.df <- cbind(sub.df,pred.causal.gene,shared.go.pathways,shared.hmdb.pathways)
  out.df$pred.causal.gene <- out.df$pred.causal.gene %>% as.character(.)
  out.df$shared.go.pathways <- out.df$shared.go.pathways %>% as.character(.)
  out.df$shared.hmdb.pathways <- out.df$shared.hmdb.pathways %>% as.character(.)
  out.df <- as.data.frame(out.df)
  out.df$shared.hmdb.pathways[out.df$shared.hmdb.pathways == "character(0)"] = "NA"

  out2.df <- c()
  for (i in 1:dim(out.df)[1]){
    sgp <- out.df$shared.go.pathways[i]
    shp <- out.df$shared.hmdb.pathways[i]
    if (sgp == "NA" & shp == "NA"){
      pass = 0 
    } else{
      out2.df <- rbind(out2.df,out.df[i,])
    }
    
  }
  return(out2.df)
}


```


```{r}

tab2.df <- make_table_2_df()

tab2.df$Tissue <- gsub("_","-",tab2.df$Tissue)

tab2.df$shared.go.pathways <- map(tab2.df$shared.go.pathways,function(s){
  vec <- strsplit(s,split=",")[[1]] 
  vec <- vec[vec !="NA"] 
  vec %>% length(.)}) %>% as.integer(.)
tab2.df$shared.hmdb.pathways <- map(tab2.df$shared.hmdb.pathways,function(s){
  vec <- strsplit(s,split=",")[[1]] 
  vec <- vec[vec !="NA"] 
  vec %>% length(.)}) %>% as.integer(.)

tab2.df <- arrange(tab2.df,metabolite)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = blues9[1:2], col=NA),
          fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="navyblue", fontface=4L)),
  rowhead=list(fg_params=list(col="white", fontface=3L)))
tab2.df$zscore <- map(tab2.df$zscore, function(n){signif(n,3)}) %>% as.numeric(.)
tab2.df$pvalue <- map(tab2.df$pvalue, function(n){signif(n,3)}) %>% as.numeric(.)
tab2.df$FDR_studywide <- map(tab2.df$FDR_studywide, function(n){signif(n,3)}) %>% as.numeric(.)

tabgrob <- tableGrob(tab2.df,theme=tt3)
grid.draw(tabgrob)
  
```


```{r}

dim(tab2.df)[1] # 57 associations involve novel genes that share the same metabolic pathway as a predicted_causal gene for that metabolite 
t <- dplyr::select(tab2.df,one_of("gene_name","metabolite")); t<-t[!duplicated(t),]
dim(t)[1] # corresponding to 24 unique gene/metabolite pairs and 
t$gene_name %>% unique(.) %>% length(.) # 24 unique genes 

```


## Re-evaluation (second part)

826 Total significant gene-metabolite pairs 

Of the 312 pairs that do not involve a predicted causal gene and are not bystander genes wrt the metabolite, 
  12 pairs (corresponding to 12 unique genes) involve a gene that is in the same metabolic pathway as a predicted causal 
  gene for that metabolite 

Therefore, upwards of 3.8% (12/312) of novel genes-metabolite are biologically plausible, lower range  


```{r}

out.df <- c() 
for (i in 1:dim(t)[1]){
  sub.df <- t[i,]
  m <- sub.df$metabolite; g <- sub.df$gene_name
  c <- filter(sub,metabolite==m,gene_name==g)$predicted_causal
  p <- filter(oneMb.df,Metabo_id==m,Gene==g)$pos_flag
  check1 <- nchar(c) > 1 # is the gene a predited causal gene for the metabolite 
  check2 <- p == "within-1MB-of-GWAS" # is gene w/n 1Mb for a GWAS signal for that metabolite 
  if (check1==FALSE & check2 == FALSE){
    out.df <- rbind(out.df,sub.df)
  }
}

filt2.df <- out.df # 12 pairs 
filt2.df %>% dim(.) # 12 novel gene-metabolite pairs (12 unique genes)
12/312  # 0.03846154
filt2.df$gene_name %>% unique(.) %>% length(.) # 12 unique genes

```




## Write updated data frames 

append to dfs 

```{r}

filt.df <- filt.df %>% dplyr::select(.,one_of("gene_name","metabolite"))
filt2.df <- filt2.df %>% dplyr::select(.,one_of("gene_name","metabolite"))


append_to_filt <- function(filt.df){
  filt.df$gene <- map(filt.df$gene_name, function(gn){filter(mtxn.df,gene_name==gn)$gene %>% 
      unique(.)}) %>% as.character(.)
  filt.df <- dplyr::select(filt.df,one_of("gene","gene_name","metabolite"))
  filt.df$metabolite_name <- map(filt.df$metabolite, function(m){filter(mtxn.df,metabolite==m)$metabolite_name %>% 
      unique(.)}) %>% as.character(.)  
  filt.df$num_sig_tiss <- map(1:dim(filt.df)[1],function(i){
    (filter(mtxn.df,gene==filt.df$gene[i],metabolite==filt.df$metabolite[i]) %>% dim(.))[1]
  }) %>% as.integer(.)
  most_sig_tiss <- c()
  zscore <- c()
  pvalue <- c()
  FDR_studywide <- c() 
  for (i in 1:dim(filt.df)[1]){
    sub <- (filter(mtxn.df,gene==filt.df$gene[i],metabolite==filt.df$metabolite[i]) %>% arrange(.,pvalue))[1,]
    most_sig_tiss <- append(most_sig_tiss,sub$Tissue)
    zscore <- append(zscore,sub$zscore)
    pvalue <- append(pvalue,sub$pvalue)
    FDR_studywide <- append(FDR_studywide,sub$FDR_studywide)     
  }
  filt.df$most_sig_tiss <- most_sig_tiss; filt.df$zscore <- zscore
  filt.df$pvalue <- pvalue; filt.df$FDR_studywide <- FDR_studywide
  return(filt.df)
}

```


```{r}

filt.df <- append_to_filt(filt.df)
filt2.df <- append_to_filt(filt2.df)

```


Append HMDB pathway information to data frame 1 


```{r}

append_hmdb_accession <- function(filt.df){
  f.df <- filt.df
  f.df$HMDB_Accession <- map(f.df$gene_name,function(g){
    filter(hmdb.df,gene_name==g)$accession
  }) %>% as.character(.)
  return(f.df)
}

```


Append shared information for data frame 2 

```{r}


append_shared_pathways <- function(filt2.df){
  f.df <- filt2.df
  pred.causal.gene <- c()
  shared.go.pathways <- c()
  shared.hmdb.pathways <- c() 
  pb <- txtProgressBar(min=0,max=dim(f.df)[1],style=3)
  for (i in 1:dim(f.df)[1]){
    setTxtProgressBar(pb,i)
    g <- f.df$gene_name[i]
    mid <- f.df$metabolite[i]
    causal.genes.df <- filter(metab.df,MetaboID==mid)
    pcg <-  causal.genes.df$predicted_causal %>% paste(.,collapse=",")
    query <- get_metab_annot(g)
    query.go <- query[[1]]
    query.hmdb <- query[[2]]
    reflists <-  ifelse(dim(causal.genes.df)[1]>0,map(1:length(causal.genes.df$predicted_causal), function(i){
      pc <- causal.genes.df$predicted_causal[i]
      metabs <- get_metab_annot(pc)
      return(metabs)
      }),NA)
    sgp <- ifelse(dim(causal.genes.df)[1]>0,map(1:length(reflists),function(i){ # 
      go.vec <- reflists[[i]][[1]]
      overlap <- go.vec[go.vec %in% query.go]
      if (length(overlap)==0){
        overlap <- NA
      }
      return(overlap)
    }) %>% unlist(.) %>% as.character(.)  %>% paste0(.,collapse=","),NA) 
    shp <- ifelse(dim(causal.genes.df)[1]>0,map(1:length(reflists),function(i){
      hmdb.vec <- reflists[[i]][[2]]
      overlap <- hmdb.vec[hmdb.vec %in% query.hmdb]
      if (length(overlap)==0){
        overlap <- NA
      }
      return(overlap)
    }) %>% unlist(.) %>% as.character(.) %>% paste0(.,collapse=","),NA)
    pred.causal.gene <- append(pred.causal.gene,pcg)
    shared.go.pathways <- append(shared.go.pathways,sgp)
    shared.hmdb.pathways <- append(shared.hmdb.pathways,shp)  
  }
  out.df <- cbind(f.df,pred.causal.gene,shared.go.pathways,shared.hmdb.pathways)
  out.df$pred.causal.gene <- out.df$pred.causal.gene %>% as.character(.)
  out.df$shared.go.pathways <- out.df$shared.go.pathways %>% as.character(.)
  out.df$shared.hmdb.pathways <- out.df$shared.hmdb.pathways %>% as.character(.)
  out.df <- as.data.frame(out.df)
  out.df$shared.hmdb.pathways[out.df$shared.hmdb.pathways == "character(0)"] = "NA"
  return(out.df)
}

```


```{r}

filt.df <- append_hmdb_accession(filt.df)
filt2.df <- append_shared_pathways(filt2.df)

write.table(filt.df,file="updated_supp_table1.txt",sep="\t",quote=F,row.names=F)
write.table(filt2.df,file="updated_supp_table2.txt",sep="\t",quote=F,row.names=F)

```


# Summary 

826 Total significant gene-metabolite pairs 

41 gene-metabolite pairs involve genes that are predicted causal for the metabolite 
473 significant gene-metabolite pairs are within 1Mb of significant metQTLs, likely to be bystanders (from Anne's section of manuscript)
312 significant gene-metabolite pairs are not bystander associations and 
    do not involve predicted causal genes for the metabolite [could also be 312 = 826 - (41+473) ]

96 pairs are outside of 1Mb windows, not predicted causal for the metabolite, and are in hmdb; 
    corresponding to 83 unique genes 
Therefore, upwards of 30.8%% (96/312) of novel genes-metabolite are biologically plausible

Of the 312 pairs that do not involve a predicted causal gene and are not bystander genes wrt the metabolite, 
  12 pairs (corresponding to 12 unique genes) involve a gene that is in the same metabolic pathway as a predicted causal 
  gene for that metabolite 

Therefore, upwards of 3.8% (12/312) of novel genes-metabolite are biologically plausible, lower range 
